---
title: "Perception of Covid-19 Measures by The Ontario Government During the Second Wave"
author: "Jin Long Cao, Brian Diep, Johnathan Tillman, Tanya Woloshansky"
date: 2020-10-07
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('tidyverse')
library(tidyverse)
```

# Executive summary

Petit Poll has been asked by the Rohan and Sam Party ('the Party') to survey British Columbia to understand whether people are likely to vote for the Party in the upcoming general election. We surveyed....

# Introduction


# Survey methodology 

- What is the population frame and sample
- What sampling methods will you use and why e.g. you could choose SRSWOR stratified etc. What are some of the statistical properties that the method brings to the table e.g. for SRSWOR you could discuss Wu Thompson Theorem 2.2. etc as appropriate
- How are you going to reach your desired respondents
- How much do you estimate this will cost
- What steps will you take to deal with non-response and how will non-response affect your survey
- How are you going to protect respondent privacy


and more...

## Survey
```{r}
#Simulating data
#We can always change the prob to fit our report
set.seed(666)
my_data <- tibble(
  Age  = sample(x = c("18-24", "25-34", "35-44", "45-54", "50-64", "65-74", "75-100"), size = 15987, replace = TRUE, prob = c(0.2, 0.225, 0.15, 0.20, .105, 0.07, 0.05)),
  
  Vulnerable_Population = sample(x = c("Yes", "No", "Not sure"), size = 15987, replace = TRUE, prob = c(0.2, 0.7, 0.1)),
  
  #Do you live in one of the following administrative/metropolitan regions
  Living = sample(x = c("Yes", "No"),  size = 15987, replace = TRUE, prob = c(0.8, 0.2)),
  
  Political_Party = sample(x = c("Progressive Conservative Party of Ontario",
                                 "Ontario Liberal Party",
                                 "Ontario New Democratic Party",
                                 "Green Party of Ontario",
                                 "No political affiliation", "Other"),
                           size = 15987, replace = TRUE, 
                           prob = c(0.2, 0.2, 0.2, 0.2, 0.1, 0.1)),
  
  Primary_Source = sample(x = c("Social Media", "Television", "Personal Research on the Internet", "Newspaper", "Family and Friends", "Other"), size = 15987, replace = TRUE, prob = c(0.3, 0.15, 0.2, 0.05, 0.25, 0.05)),
  
  Knowledgeable = sample(c(1:5), size = 15987, replace = TRUE, prob = c(0.25, 0.15, 0.1, 0.2, 0.3)),
  
  Mask_Frequency = sample(x = c("Always", "Most of the time", "Sometimes", "Rarely", "Never"), size = 15987, replace = TRUE, prob = c(0.25, 0.3, 0.2, 0.15, 0.1)),
  
  Mask_Type = sample(x = c("Surgical Mask", "N95 Mask", "Cloth Mask", "Other"), size = 15987, replace = TRUE, prob = c(0.35, 0.25, 0.3, 0.1)),
  
  Social_Bubble = rpois(15987, 10),
  
  Follow_Regulation = sample(x = c("Yes, I follow all store requirements",
                                   "Mostly",
                                   "Only ones I think will be enforced",
                                   "None"),
                             size = 15987, replace = TRUE, 
                           prob = c(0.25, 0.25, 0.25, 0.25)),
  Visit_non_essentials = rpois(15987, 3),
  
  Travel_Within_Canada = sample(x = c("Yes", "No"), size = 15987, replace = TRUE, prob = c(0.7066, 0.2934)), 
  
  Travel_Out_Of_Canada = sample(x = c("Yes", "No"), size = 15987, replace = TRUE, prob = c(0.98743, 0.01257)),
  
  Lockdown = sample(x = c("Strongly agree", "Slightly agree", "Neutral", "Slightly disagree", "Strongly disagree"), size = 15987, replace = TRUE, prob = c(0.20, 0.20, 0.2, 0.2, 0.2)),
  
  Travel_Border_Restriction = sample(x = c("Yes", "No"), size = 15987, replace = TRUE, prob = c(0.98743, 0.01257))
  
)

my_data %>% 
  ggplot(aes(x = Travel_Within_Canada)) +
  geom_bar()

my_data %>% 
  ggplot(aes(x=Age))+
  geom_bar()


```

Add a link to the survey and screen shots of the questions.



## Results


# Discussion 

## Weakness and areas for future


# Appendix


# References

- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open
  Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
- R Core Team (2020). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria.
  URL https://www.R-project.org/.
